---
title: "SARS-CoV-2 Wastewater Surveillance by US Region, March 2020 - December 2023"
author: Ben Kessler, Jiawen Shi, Natnael Kassahun, Steph Ponce, Suprithi Pingle
document: html_document
---
# **Introduction**
This dataset contains information about the complete timeline of SARS-CoV-2 concentrations in waste water (as of February 25, 2024)

# **Aim**
The primary objective of this analysis is to report the descriptive epidemiology of the waste water prevalence of COVID-19 by user-specified region(s) of the US.

```{r, include=FALSE}

knitr::opts_chunk$set(echo=FALSE)
here::i_am("midterm_project_report.rmd")
pacman::p_load(config)

```

```{r, data_source_parameter}

#data source
Sys.setenv(R_CONFIG_ACTIVE = "data_source")
config_list <- config::get("original_data")
data_source <- ifelse(config_list[[1]] == TRUE, "original", "updated")

```

These data come from the `r data_source` data set.

The following regions were included in this analysis:

```{r, region_paramter}

#US regions to include
Sys.setenv(R_CONFIG_ACTIVE = "default")
config_list <- config::get()
region_list <- c("Midwest", "Northeast", "South", "West")
regions_to_include <- NULL

for (i in 1:4) {
  if (config_list[[i]] == TRUE) {
    regions_to_include[i] <- region_list[[i]]
  }
  else {
    next
  }
}

print(regions_to_include)

```

**Table 1.**  SARS-CoV-2 virus level mean percentile by US region from 2020-2023.

```{r, table_1}

readRDS(here::here("output/desc_table.Rds"))

```

**Figure 1.**  SARS-Cov-2 virus level mean percentile by state in 2023.  

```{r, figure_1}
knitr::include_graphics(
  here::here("output/us_plot.png")
  )
```

**Figure 2.**  Comparison of US region and total US SARS-Cov-2 virus level mean percentile for 2023.

```{r, figure_2}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("output/bargraph.png")
  )
```

**Figure 3.** Change in SARS-Cov-2 virus level mean percentile by US region from 2020-2023.

```{r, figure_3}
knitr::include_graphics(
  here::here("output/figure_three.png")
  )
```